---
title: "R Notebook"
output: html_notebook
---

#### Load frequently used packages
```{r}
library(openxlsx)
library(ggrepel)
library(dplyr)
library(tidyr)
library(data.table)
library(broom)
library(broomExtra)
library(tibble)
library(sjstats)
library(car)
library(lme4)
library(lmerTest)
library(ggplot2)
library(tibble)
library(modelr)
library(tidyverse)
#library(miceadds)
library(ggforce)
require(openxlsx)
library(tidyverse)
library(caret)
library(glmnet)
library(ggplot2)
select <- dplyr::select
```

#### Print session info for reproducibility
```{r}
sessionInfo()
```

#### Define in/out directory
```{r}
dir <- "/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/Celiac_MultiOmics/GRS/DSMIG_Shared/Manuscript_Figure1/Data"
```

#### Put a copy of the fam file and bim file in the shared manuscript directory
```{bash}
cp -a '/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/MEGA/Plink/Espinosa_MEGA2_HTP_GS_08132019_updated_callrate_passing_QC.fam' \
'/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/Celiac_MultiOmics/GRS/DSMIG_Shared/Manuscript_Figure1/Data'

cp -a '/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/MEGA/Plink/Espinosa_MEGA2_HTP_GS_08132019_updated_callrate_passing_QC.bim' \
'/Users/shawjes/Dropbox/EspinosaGroup/ANALYSIS/Celiac_MultiOmics/GRS/DSMIG_Shared/Manuscript_Figure1/Data'
```

#### Check if the plink files were successfully copied into the shared directory
```{r}
list.files(dir) %>%
  as.data.frame() %>%
  `colnames<-`("filename") %>%
  filter(grepl("bin", filename)==TRUE | grepl("fam", filename)==TRUE)
```

#### Read Plink *.bim file
```{r}
setwd(dir)
MEGA.fam <- fread("Espinosa_MEGA2_HTP_GS_08132019_updated_callrate_passing_QC.fam") %>%
  select(V1, V2) %>%
  rename(MEGA.FID = V1, MEGA.IID = V2)

MEGA.fam
```

#### Read in comorb data and produce a key mapping HTP FamilyID, RecordID, and LabIDs
```{r}
setwd("/Users/shawjes/Dropbox/EspinosaGroup/DATA_MAIN/META/OneDrive_1_2-17-2022")
IDs.022021 <- fread("HTP_CLEANED_02_2021_v0.5_COMORB_MASTER_RecordID_vs_LabID.Labels.tsv") %>%
  select(FamilyID, RecordID, LabID) %>%
  filter(!is.na(LabID)) %>%
  unique()

IDs.022021
```

#### Test out parsing of MEGA IIDs to extract LabID used for genotyping
```{r}
MEGA.fam %>%
  separate(MEGA.IID, into = c("a", "b", "c"), sep = "_", extra = "merge", remove = FALSE)

MEGA.fam %>%
  separate(MEGA.IID, into = c("a", "b", "c"), sep = "_", extra = "merge", remove = FALSE) %>%
  filter(grepl("HTP", c)==FALSE & grepl("NA", c)==FALSE)

print("Looks good, all rows in column c contain the 'HTP' prefix that indicates a participant sample.")
```

#### Produce an ID key that shows standard HTP/P4C identifiers alongside FID and IID that appear in the MEGA2 data received from CCPM
```{r}
MEGA.IDkey <- MEGA.fam %>%
  separate(MEGA.IID, into = c("a", "b", "c"), sep = "_", extra = "merge", remove = FALSE) %>%
  select(-c(a, b)) %>%
  rename(MEGA.LabID = c) %>%
  mutate(HTP_participant = ifelse(grepl("HTP", MEGA.LabID)==TRUE, "Yes", "No")) %>%
  mutate(In_plink_fam = 1) %>%
  left_join(IDs.022021, by = c("MEGA.LabID" = "LabID")) %>%
  select(FamilyID, RecordID, MEGA.FID, MEGA.IID, MEGA.LabID, HTP_participant, In_plink_fam) %>%
  unique()

MEGA.IDkey
```

#### Save the ID key as a TSV and as a CSV
```{r}
setwd(dir)
fwrite(MEGA.IDkey, "MEGA_041822_MEGA2_to_HTP_ID_key_v0.1_JRS.csv")
fwrite(MEGA.IDkey, "MEGA_041822_MEGA2_to_HTP_ID_key_v0.1_JRS.tsv", sep = "\t")
```
